Think deep for reef resilience

### 2025-08-07 paper/paper_draft.qmd
Remove supplementary figures that are just from extra GCMs
Only ACCESS-ESM1-5 has individual results displayed in the main paper text or supplementary results.
This intermediate GCM's results are indicative of other GCMs.
GCM comparison figures of clustering are still included in paper content.

: Table 1: Information for each GCM used in this investigation including ECS value,
interpreted relative severity and where results are individual GCM results are displayed.
GCMs in context with assessed plausible ranges of ECS values are displayed in [@fig-ecs].
ACCESS-ESM1-5 is taken as an intermediate model based on the range of ECS values, thus
results are displayed in the main text. ECS value sources: .

@@

Across many bioregions timeseries clustering identified a ‘low’ cluster containing reefs
that suffer larger declines and generally have lower levels of coral cover than the other
clusters, a ‘medium’ cluster with intermediate levels of coral cover, and a ‘high’ cluster
with greater levels of cover recovery, with persistence to 2050 in some bioregions
(@sfig-access-esm1-5-bioregion-timeseries; @sfig-access-cm2-bioregion-timeseries;
@sfig-ec-earth3-veg-bioregion-timeseries; @sfig-gfdl-cm4-bioregion-timeseries;
@sfig-noresm2-mm-bioregion-timeseries).

In many bioregions, median reef depth decreases as cluster coral cover decreases, with reefs
within the ‘high’ cluster commonly having the greatest median reef depths within the
bioregion (@fig-access-esm1-5-bioregion-depth; @sfig-access-cm2-bioregion-depth;
@sfig-ec-earth3-veg-bioregion-depth; @sfig-gfdl-cm4-bioregion-depth;
@sfig-noresm2-mm-bioregion-depth).

@@

In many bioregions reefs in the 'low' cover cluster have the lowest cover - DHW timeseries
correlation levels, while reefs in 'high' clusters have the highest correlation levels
(@sfig-access-esm1-5-bioregion-dhwcor; @sfig-access-cm2-bioregion-dhwcor;
@sfig-ec-earth3-veg-bioregion-dhwcor; @sfig-gfdl-cm4-bioregion-dhwcor;
@sfig-noresm2-mm-bioregion-dhwcor).

Conversely, there are no consistent patterns across bioregions in reef area-relative cover
and mean timeseries DHW levels, with clusters having similar mean DHW levels in many
bioregions (@sfig-access-esm1-5-bioregion-dhwmean; @sfig-access-cm2-bioregion-dhwmean;
@sfig-ec-earth3-veg-bioregion-dhwmean; @sfig-gfdl-cm4-bioregion-dhwmean;
@sfig-noresm2-mm-bioregion-dhwmean).

Similarly, there are no substantial differences between reef clusters in mean thermal
tolerance levels (@sfig-access-esm1-5-bioregion-dhwtol; @sfig-access-cm2-bioregion-dhwtol;
@sfig-ec-earth3-veg-bioregion-dhwtol; @sfig-gfdl-cm4-bioregion-dhwtol;
@sfig-noresm2-mm-bioregion-dhwtol).

There are no consistent differences across bioregions and GCMs between clusters in their
reefs' initial proportion of habitable area occupied by corals
(@sfig-access-esm1-5-bioregion-initialproportion;
@sfig-access-cm2-bioregion-initialproportion;
@sfig-ec-earth3-veg-bioregion-initialproportion; @sfig-gfdl-cm4-bioregion-initialproportion;
@sfig-noresm2-mm-bioregion-initialproportion).

Across most bioregions there are not consistent differences between clusters in connectivity
metrics such as source-to-sink connectivity ratio (@sfig-access-esm1-5-bioregion-logsosi;
@sfig-access-cm2-bioregion-logsosi; @sfig-ec-earth3-veg-bioregion-logsosi;
@sfig-gfdl-cm4-bioregion-logsosi; @sfig-noresm2-mm-bioregion-logsosi) and total connectivity
strength (@sfig-access-esm1-5-bioregion-totalconnectivity;
@sfig-access-cm2-bioregion-totalconnectivity;
@sfig-ec-earth3-veg-bioregion-totalconnectivity; @sfig-gfdl-cm4-bioregion-totalconnectivity;
@sfig-noresm2-mm-bioregion-totalconnectivity).

@@

Our findings on the importance of connectivity may differ due to the highly uncertain nature
of larval connectivity modelling [@vogt-vincent2023], the way larval flows and juvenile
recruitment are implemented in ADRIA - CoralBlox, and the magnitude of early bleaching
events decreasing the breeding capacity of strongly connected reefs. Analysing daily
connectivity data Vogt-Vincent et al. [-@vogt-vincent2023] suggest that time-mean
connectivity may not be suitable for capturing the seasonal - ecological time scale
connectivity between locations due to high daily variability in connectivity data. This is
supported by the connectivity data used in our investigation, which have large relative
standard deviations for connectivity between reefs over the 24 days used
[@sfig-connectivity-rsd]. Using alternative summarisation methods that capture
ecological-timescale patterns in coral larval dispersal, such as cumulated implicit
connectivity [@vogt-vincent2023], future investigations may find increased importance of
both depth and reef connectivity for resilience, in line with other investigations
[@hock2017; @torres2024]. Further improvements and research in connectivity modelling are
necessary to provide a robust comparison of all factors influencing reef resilience.

@@

### Bioregion scale

@@

## ACCESS-CM2

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
ACCESS-CM2.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/depth_carbonate_scatter.png){#sfig-access-cm2-depth-carbonate
width="14.82cm"}

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef log (natural)
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. Point colour represents log (natural) scaled reef total connectivity strength.
Displays results from median timeseries using
ACCESS-CM2.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/log_total_strength_carbonate_scatter.png){#sfig-access-cm2-conn-carbonate
width="14.82cm"}

### Bioregion scale

![Reef area-relative coral cover timeseries clustered within each bioregion. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2040 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2040 within each
bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5
to 97.5 percentile confidence intervals. Displays results for ACCESS-CM2
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/bioregion_cover_timeseries.png){#sfig-access-cm2-bioregion-timeseries
width="14.82cm"}

![Median depth for reefs within each bioregion, separated into clusters assigned by
timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median
area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange)
are clusters with intermediate median area-relative cover between 2030 and 2040 within each
bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover
between 2030 and 2040 within each bioregion. Solid lines represent the median depth for each
cluster. Displays results for ACCESS-CM2
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/depth_bioregion_violin.png){#sfig-access-cm2-bioregion-depth
width="14.82cm"}

![Mean timeseries DHW values for reefs within each bioregion, separated into clusters
assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest
median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters
(orange) are clusters with intermediate median area-relative cover between 2030 and 2040
within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW level for each cluster. Displays results for ACCESS-CM2
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/mean_dhw_bioregion_violin.png){#sfig-access-cm2-bioregion-dhwmean
width="14.82cm"}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median cover – DHW correlation level for each cluster. Displays results
for ACCESS-CM2
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-access-cm2-bioregion-dhwcor
width="14.82cm"}

![Mean timeseries DHW tolerance value for reefs within each bioregion, separated into
clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the
lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’
clusters (orange) are clusters with intermediate median area-relative cover between 2030 and
2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW tolerance value for each cluster. Displays results for ACCESS-CM2
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/dhw_tolerance_bioregion_violin.png){#sfig-access-cm2-bioregion-dhwtol
width="14.82cm"}

![Initial proportion of coral habitable area occupied by corals for reefs within each
bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green)
are clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median proportion of habitable area occupied for each cluster. Displays
results for ACCESS-CM2
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/initial_proportion_bioregion_violin.png){#sfig-access-cm2-bioregion-initialproportion
width="14.82cm"}

![Log (natural) source – to – sink connectivity ratios for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines within each violin represent the median log (natural) source – to – sink ratio for
each cluster and grey points represent the ratio value for each reef. Vertical grey lines
represent a log (natural) source – to – sink ratio of 0 indicating equal outgoing and
incoming connectivity levels. Displays results for ACCESS-CM2
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/so_to_si_bioregion_violin.png){#sfig-access-cm2-bioregion-logsosi
width="14.82cm"}

![Log (natural) total connectivity strength for reefs within each bioregion, separated into
clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the
lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’
clusters (orange) are clusters with intermediate median area-relative cover between 2030 and
2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines within each
violin represent the median log (natural) total connectivity strength for each cluster.
Displays results for ACCESS-CM2
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-CM2/bioregion/log_total_strength_bioregion_violin.png){#sfig-access-cm2-bioregion-totalconnectivity
width="14.82cm"}

## EC-Earth3-Veg

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
EC-Earth3-Veg.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/depth_carbonate_scatter.png){#sfig-ec-earth3-veg-depth-carbonate
width="14.82cm"}

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef log (natural)
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. Point colour represents log (natural) scaled reef total connectivity strength.
Displays results from median timeseries using
EC-Earth3-Veg.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/log_total_strength_carbonate_scatter.png){#sfig-ec-earth3-veg-conn-carbonate
width="14.82cm"}

### Bioregion scale

![Reef area-relative coral cover timeseries clustered within each bioregion. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2040 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2040 within each
bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5
to 97.5 percentile confidence intervals. Displays results for EC-Earth3-Veg
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/bioregion/bioregion_cover_timeseries.png){#sfig-ec-earth3-veg-bioregion-timeseries
width="14.82cm"}

![Median depth for reefs within each bioregion, separated into clusters assigned by
timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median
area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange)
are clusters with intermediate median area-relative cover between 2030 and 2040 within each
bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover
between 2030 and 2040 within each bioregion. Solid lines represent the median depth for each
cluster. Displays results for EC-Earth3-Veg
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/bioregion/depth_bioregion_violin.png){#sfig-ec-earth3-veg-bioregion-depth
width="14.82cm"}

![Mean timeseries DHW values for reefs within each bioregion, separated into clusters
assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest
median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters
(orange) are clusters with intermediate median area-relative cover between 2030 and 2040
within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW level for each cluster. Displays results for EC-Earth3-Veg
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/bioregion/mean_dhw_bioregion_violin.png){#sfig-ec-earth3-veg-bioregion-dhwmean
width="14.82cm"}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median cover – DHW correlation level for each cluster. Displays results
for EC-Earth3-Veg
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-ec-earth3-veg-bioregion-dhwcor
width="14.82cm"}

![Mean timeseries DHW tolerance value for reefs within each bioregion, separated into
clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the
lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’
clusters (orange) are clusters with intermediate median area-relative cover between 2030 and
2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW tolerance value for each cluster. Displays results for EC-Earth3-Veg
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/bioregion/dhw_tolerance_bioregion_violin.png){#sfig-ec-earth3-veg-bioregion-dhwtol
width="14.82cm"}

![Initial proportion of coral habitable area occupied by corals for reefs within each
bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green)
are clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median proportion of habitable area occupied for each cluster. Displays
results for EC-Earth3-Veg
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/bioregion/initial_proportion_bioregion_violin.png){#sfig-ec-earth3-veg-bioregion-initialproportion
width="14.82cm"}

![Log (natural) source – to – sink connectivity ratios for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines within each violin represent the median log (natural) source – to – sink ratio for
each cluster and grey points represent the ratio value for each reef. Vertical grey lines
represent a log (natural) source – to – sink ratio of 0 indicating equal outgoing and
incoming connectivity levels. Displays results for EC-Earth3-Veg
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/bioregion/so_to_si_bioregion_violin.png){#sfig-ec-earth3-veg-bioregion-logsosi
width="14.82cm"}

![Log (natural) total connectivity strength for reefs within each bioregion, separated into
clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the
lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’
clusters (orange) are clusters with intermediate median area-relative cover between 2030 and
2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines within each
violin represent the median log (natural) total connectivity strength for each cluster.
Displays results for EC-Earth3-Veg
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/EC-Earth3-Veg/bioregion/log_total_strength_bioregion_violin.png){#sfig-ec-earth3-veg-bioregion-totalconnectivity
width="14.82cm"}

## GFDL-CM4

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
GFDL-CM4.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/depth_carbonate_scatter.png){#sfig-gfdl-cm4-depth-carbonate
width="14.82cm"}

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef log (natural)
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. Point colour represents log (natural) scaled reef total connectivity strength.
Displays results from median timeseries using
GFDL-CM4.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/log_total_strength_carbonate_scatter.png){#sfig-gfdl-cm4-conn-carbonate
width="14.82cm"}

### Bioregion scale

![Reef area-relative coral cover timeseries clustered within each bioregion. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2040 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2040 within each
bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5
to 97.5 percentile confidence intervals. Displays results for GFDL-CM4
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/bioregion/bioregion_cover_timeseries.png){#sfig-gfdl-cm4-bioregion-timeseries
width="14.82cm"}

![Median depth for reefs within each bioregion, separated into clusters assigned by
timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median
area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange)
are clusters with intermediate median area-relative cover between 2030 and 2040 within each
bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover
between 2030 and 2040 within each bioregion. Solid lines represent the median depth for each
cluster. Displays results for GFDL-CM4
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/bioregion/depth_bioregion_violin.png){#sfig-gfdl-cm4-bioregion-depth
width="14.82cm"}

![Mean timeseries DHW values for reefs within each bioregion, separated into clusters
assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest
median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters
(orange) are clusters with intermediate median area-relative cover between 2030 and 2040
within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW level for each cluster. Displays results for GFDL-CM4
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/bioregion/mean_dhw_bioregion_violin.png){#sfig-gfdl-cm4-bioregion-dhwmean
width="14.82cm"}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median cover – DHW correlation level for each cluster. Displays results
for GFDL-CM4
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-gfdl-cm4-bioregion-dhwcor
width="14.82cm"}

![Mean timeseries DHW tolerance value for reefs within each bioregion, separated into
clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the
lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’
clusters (orange) are clusters with intermediate median area-relative cover between 2030 and
2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW tolerance value for each cluster. Displays results for GFDL-CM4
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/bioregion/dhw_tolerance_bioregion_violin.png){#sfig-gfdl-cm4-bioregion-dhwtol
width="14.82cm"}

![Initial proportion of coral habitable area occupied by corals for reefs within each
bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green)
are clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median proportion of habitable area occupied for each cluster. Displays
results for GFDL-CM4
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/bioregion/initial_proportion_bioregion_violin.png){#sfig-gfdl-cm4-bioregion-initialproportion
width="14.82cm"}

![Log (natural) source – to – sink connectivity ratios for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines within each violin represent the median log (natural) source – to – sink ratio for
each cluster and grey points represent the ratio value for each reef. Vertical grey lines
represent a log (natural) source – to – sink ratio of 0 indicating equal outgoing and
incoming connectivity levels. Displays results for GFDL-CM4
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/bioregion/so_to_si_bioregion_violin.png){#sfig-gfdl-cm4-bioregion-logsosi
width="14.82cm"}

![Log (natural) total connectivity strength for reefs within each bioregion, separated into
clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the
lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’
clusters (orange) are clusters with intermediate median area-relative cover between 2030 and
2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines within each
violin represent the median log (natural) total connectivity strength for each cluster.
Displays results for GFDL-CM4
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/GFDL-CM4/bioregion/log_total_strength_bioregion_violin.png){#sfig-gfdl-cm4-bioregion-totalconnectivity
width="14.82cm"}

## NorESM2-MM

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
NorESM2-MM.](../outputs/ADRIA_results/HighResCoralStress/figs/NorESM2-MM/depth_carbonate_scatter.png){#sfig-noresm2-mm-depth-carbonate
width="14.82cm"}

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2099. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef log (natural)
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. Point colour represents log (natural) scaled reef total connectivity strength.
Displays results from median timeseries using
NorESM2-MM.](../outputs/ADRIA_results/HighResCoralStress/figs/NorESM2-MM/log_total_strength_carbonate_scatter.png){#sfig-noresm2-mm-conn-carbonate
width="14.82cm"}

### Bioregion scale

![Reef area-relative coral cover timeseries clustered within each bioregion. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2040 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2040 within each
bioregion. Solid lines represent the cluster median timeseries and bands represent the 2.5
to 97.5 percentile confidence intervals. Displays results for NorESM2-MM
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/NorESM2-MM/bioregion/bioregion_cover_timeseries.png){#sfig-noresm2-mm-bioregion-timeseries
width="14.82cm"}

![Median depth for reefs within each bioregion, separated into clusters assigned by
timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median
area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange)
are clusters with intermediate median area-relative cover between 2030 and 2040 within each
bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover
between 2030 and 2040 within each bioregion. Solid lines represent the median depth for each
cluster. Displays results for NorESM2-MM
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/NorESM2-MM/bioregion/depth_bioregion_violin.png){#sfig-noresm2-mm-bioregion-depth
width="14.82cm"}

![Mean timeseries DHW values for reefs within each bioregion, separated into clusters
assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest
median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’ clusters
(orange) are clusters with intermediate median area-relative cover between 2030 and 2040
within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW level for each cluster. Displays results for NorESM2-MM
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/NorESM2-MM/bioregion/mean_dhw_bioregion_violin.png){#sfig-noresm2-mm-bioregion-dhwmean
width="14.82cm"}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2040 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2040 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2040 within each bioregion. Solid
lines represent the median cover – DHW correlation level for each cluster. Displays results
for NorESM2-MM
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/NorESM2-MM/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-noresm2-mm-bioregion-dhwcor
width="14.82cm"}

![Mean timeseries DHW tolerance value for reefs within each bioregion, separated into
clusters assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the
lowest median area-relative cover between 2030 and 2040 within each bioregion. ‘Medium’
clusters (orange) are clusters with intermediate median area-relative cover between 2030 and
2040 within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2040 within each bioregion. Solid lines represent the
median DHW tolerance value for each cluster. Displays results for NorESM2-MM
GCM.](../outputs/ADRIA_results/HighResCoralStress/figs/NorESM2-MM/bioregion/dhw_tolerance_bioregion_violin.png){#sfig-noresm2-mm-bioregion-dhwtol
width="14.82cm"}

>>>>

: Table 1: Information for each GCM used in this investigation including ECS value,
interpreted relative severity and where results are individual GCM results are displayed.
GCMs in context with assessed plausible ranges of ECS values are displayed in [@fig-ecs].
ACCESS-ESM1-5 is taken as an intermediate model based on the range of ECS values, when
assessing factors impacting reef resilience these results are indicative of results from all
GCMs assessed. Therefore, inidividual results for other GCMs are available at **insert repo
citation**. ECS value sources: .

@@

Across many bioregions timeseries clustering identified a ‘low’ cluster containing reefs
that suffer larger declines and generally have lower levels of coral cover than the other
clusters, a ‘medium’ cluster with intermediate levels of coral cover, and a ‘high’ cluster
with greater levels of cover recovery, with persistence to 2050 in some bioregions
(@sfig-access-esm1-5-bioregion-timeseries).

In many bioregions, median reef depth decreases as cluster coral cover decreases, with reefs
within the ‘high’ cluster commonly having the greatest median reef depths within the
bioregion (@fig-access-esm1-5-bioregion-depth).

@@

In many bioregions reefs in the 'low' cover cluster have the lowest cover - DHW timeseries
correlation levels, while reefs in 'high' clusters have the highest correlation levels
(@sfig-access-esm1-5-bioregion-dhwcor).

Conversely, there are no consistent patterns across bioregions in reef area-relative cover
and mean timeseries DHW levels, with clusters having similar mean DHW levels in many
bioregions (@sfig-access-esm1-5-bioregion-dhwmean).

Similarly, there are no substantial differences between reef clusters in mean thermal
tolerance levels (@sfig-access-esm1-5-bioregion-dhwtol).

There are no consistent differences across bioregions and GCMs between clusters in their
reefs' initial proportion of habitable area occupied by corals
(@sfig-access-esm1-5-bioregion-initialproportion).

Across most bioregions there are not consistent differences between clusters in connectivity
metrics such as source-to-sink connectivity ratio (@sfig-access-esm1-5-bioregion-logsosi)
and total connectivity strength (@sfig-access-esm1-5-bioregion-totalconnectivity).

@@

Our findings on the importance of connectivity may differ due to the highly uncertain nature
of larval connectivity modelling [@vogt-vincent2023], the way larval flows and juvenile
recruitment are implemented in ADRIA - CoralBlox, and the magnitude of early bleaching
events decreasing the breeding capacity of strongly connected reefs. Analysing daily
connectivity data Vogt-Vincent et al. [-@vogt-vincent2023] suggest that time-mean
connectivity may not be suitable for capturing the seasonal - ecological time scale
connectivity between locations due to high daily variability in connectivity data. This is
supported by the connectivity data used in our investigation, which have large relative
standard deviations for connectivity between reefs over the 24 days used
(@sfig-connectivity-rsd). Using alternative summarisation methods that capture
ecological-timescale patterns in coral larval dispersal, such as cumulated implicit
connectivity [@vogt-vincent2023], future investigations may find increased importance of
both depth and reef connectivity for resilience, in line with other investigations
[@hock2017; @torres2024]. Further improvements and research in connectivity modelling are
necessary to provide a robust comparison of all factors influencing reef resilience.

@@

### Bioregion scale clustering

@@
